---
title: "Assignment 3: parallel computing"
author: "Stijn Hanssen"
date: "Oktober 2018"
output: html_document
---

### Clear working space
First step is to make sure that the working space is empty.
```{r}
# clear all variables
rm(list=ls()) 
```

### Set path working directory
The directory path is set to the stored dataset.
```{r}
# set working directory
setwd("C:/Users/stijn/Desktop/Master systems biology/Systems biology year 2/scientific computing/parallelprocessing") 
```

### Load packages
Setting the size memory for the project and loading the required packages for specific functions. If the packaged aren't present they will be automatically be downloaded.
```{r}
# setting working memory for this job to 8gb
options(java.parameters = "-Xmx8g")
# install pacman
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
# p_load: instal and load the packages
# ittertools: for proper nextElem support
# rcdk: acces functioality of jave framework
# doParallel: provides parallel backend for the dopar function
pacman::p_load("parallel","rcdk","itertools","doParallel")
```

### Import data
The `aid624202.sdf` will be loaded to compute the descriptor matrix. In the `while` loop an iteration will take place which will add the molecules one by one to the `mols` list. This will be done until the last molecule of the subset is submitted to `mols`.
```{r}
# total amount molecules in sdf file
max_n_molecules = 377550
# amount molecules imported
n = 1000
# iload.molecules: iterating version of loader
iter <- iload.molecules('C:/Users/stijn/Desktop/Master systems biology/Systems biology year 2/scientific computing/parallelprocessing/aid624202.sdf', type='sdf')
# text progress bar
pb <- txtProgressBar(min = 0, max = n , style = 3)
# initiate value
step = 0 
# set a list
mols = list()
# while loop to iterate until n is reached
while(hasNext(iter)) {
step = step + 1
# generic function to prodcue values, save values in mols    
    mols[[step]] <- nextElem(iter)
setTxtProgressBar(pb,step)
# break the iteration when the last molecules has been imported
if(step == n){
    break
  }
}
close(pb)
```

### List descriptors
The list of descriptors wanted to calculate are depicted in this line of R-code
```{r}
# return the descriptor categories that are available
descNames <- unique(unlist(sapply(get.desc.categories()[2:4], get.desc.names)))
```

### Convert into smiles
In order to make the parallel computing work, we converted the molecules into smiles. 
```{r}
# produce vector containing the atomic components
mols = unlist(mols)
# function generate smiels representation of IAtomContainer object
smiles = lapply(mols,get.smiles)
```
